---
title: "01_Statistics of Returns"
author: "Richard Bui"
format: html
editor: visual
---

## Step 1. Load libraries

```{r}
# before loading, you need to install each package with install.packages("NAME_OF_PACKAGE")
library(tidyverse) # for MacOS: library(dplyr)
library(haven)
library(skimr)
```

## Step 2. Load data

Normally, we import data from a source file such as `csv` , Excel (`xls` ), Stata (`dta`), or SAS (`sas7bdat`). In R, we use corresponding command to load the data into R workspace.

For example, we read `csv` with `read_csv`, read `dta` with `read_dta` (from `haven` package) as an R object with the name `Ret`. So, it looks like `Ret = read_dta("FILE_NAME")` as follows.

```{r}
Ret = read_dta(file = "data/Ret.dta")
```

Look at the `Environment` tab in RStudio to see the new object `Ret`. Click it to view it.

**Practice**

> Can you import market return data from `data/Mret.dta` ?

## Step 3. Statistics of `Ret` data

We will work as Q&A in this part.

Q1. How many observations in data?

```{r}
dim(Ret) # dim = dimension
```

Q2. How many years of data?

```{r}

```

Q3. How many firms in the data?

```{r}

```

Q4. What is the mean and SD of returns over sample?

```{r}

```

Q5. Calculate the mean of returns over year?

```{r}

```

Q6. Draw a line plot of mean returns over year?

```{r}

```
