---
title: "04 Fama-French regression and alphas"
author: "Richard Bui"
format: html
editor: visual
---

## Step 1. Load libraries

```{r}
# before loading, you need to install each package with install.packages("NAME_OF_PACKAGE")
library(tidyverse) # for MacOS: library(dplyr)
library(haven)
library(skimr)

```

## Step 2. Re-create the portfolio returns as Lecture 3

Load data on returns and firm factors:

```{r}
Ret = read_dta(file = "data/Ret.dta")
Size = read_dta(file = "data/Size.dta")
Beta = read_dta(file = "data/Beta.dta")
```

Merge data:

```{r}
Data = Ret %>% 
  inner_join(Size, by=c("permno", "yyyymm")) %>% 
  inner_join(Beta, by=c("permno", "yyyymm"))

Data = Data %>% 
  mutate(size = ifelse(me>0, log(me), NA))

Data = Data %>% 
  filter(year(date)>=1990)

Data2 = Data %>% 
  group_by(permno) %>% 
  arrange(permno, yyyymm) %>% 
  mutate(size_l1 = lag(size),
         beta_l1 = lag(beta)) %>% 
  ungroup()
```

Equal weighted returns:

```{r}
PortfolioReturn = Data2 %>% 
  mutate(g_size = ntile(size_l1, n=10), 
         .by = date) %>%
  summarize(p_ret = mean(ret, na.rm=TRUE), 
            .by = c(date, g_size)) 
```

Then, we have the time series of returns for 10 portfolios:

```{r}
PortfolioReturn
```

## Step 3. Load Fama-French 3-factor returns

```{r}
Ff3 = read_csv("data/F-F_Research_Data_Factors.CSV")
```

We observe that Ff3 data has a numeric column `yyyymm` but PortfolioReturn data has a date-type `date` column. We can't merge using these two columns since they are not in the same data type.

Thus, we need to create a `yyyymm` column in the `PortfolioReturn` data (note that we also can do oppositely, create a `date` column in Ff3, which I leave to you as self homework):

```{r}
PortfolioReturn %>% 
  # create yyyymm
  mutate(yyyymm = as.character(date)) %>% 
  mutate(yyyymm = str_sub(yyyymm, 1, 7)) %>% 
  mutate(yyyymm = str_replace(yyyymm, "-", "")) %>% 
  mutate(yyyymm = as.numeric(yyyymm)) 
```

I will analyze step by step so that you can understand in class.

```{r}
PortfolioReturn = PortfolioReturn %>% 
  # create yyyymm
  mutate(yyyymm = as.character(date)) %>% 
  mutate(yyyymm = str_sub(yyyymm, 1, 7)) %>% 
  mutate(yyyymm = str_replace(yyyymm, "-", "")) %>% 
  mutate(yyyymm = as.numeric(yyyymm)) %>% 
  # merge with Ff3
  inner_join(Ff3, by="yyyymm")
```

Next, we need to take portfolio return to minus risk-free rates to get excess returns (after risk free rates):

```{r}
PortfolioReturn = PortfolioReturn %>% 
  mutate(p_eret = p_ret - rf)
```

## Step 4: Run regression by each portfolio to get alphas

```{r}
PortfolioReturn %>% count(g_size)
```

We have 10 portfolios based on firm size.

We need to run each regression by each group. So totally, we will run 10 regressions.

**Portfolio 1: Smallest firms**

```{r}
PortfolioReturn %>% 
  filter(g_size==1) %>% 
  lm(p_eret ~ mkt_rf + smb + hml, data=.) %>% 
  summary()
```

In this regression, the smallest size portfolio has the alpha of 1.10 and it's significant.

**Portfolio 10: Biggest firms**

```{r}
PortfolioReturn %>% 
  filter(g_size==10) %>% 
  lm(p_eret ~ mkt_rf + smb + hml, data=.) %>% 
  summary()
```

The biggest firms have the alphas of -0.02 and it is not significant.

**How about the Long-Short portfolio?**

```{r}
PortfolioReturn %>% 
  filter(g_size==10) %>% # Largest firms
  rename(p10_ret = p_ret)

PortfolioReturn %>% 
  filter(g_size==1) %>% # Smallest firms
  select(date, p1_ret = p_ret)

```

Merge above two data:

```{r}
LongShort = inner_join(
  PortfolioReturn %>% 
  filter(g_size==10) %>% # Largest firms
  rename(p10_ret = p_ret),

  PortfolioReturn %>% 
  filter(g_size==1) %>% # Smallest firms
  select(date, p1_ret = p_ret),

  by="date"
)
```

Then, we calculate the long-short return by take `p1_ret` minus `p10_ret` :

```{r}
LongShort = LongShort %>% 
  mutate(ls_ret = p1_ret - p10_ret)
```

Run regression to get alphas:

```{r}
lm(ls_ret ~ mkt_rf + smb + hml, data=LongShort) %>% 
  summary()
```

Long-short alpha is 1.12586 and it's very significant, confirming the size premium in the stock market AFTER adjusted by the Fama-French three factor model (Fama & French, 1992, 1996).
